---
title: "Getting Started with `netjack`"
author: "Teague Henry"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{netjackIntro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette provides an introduction to the `netjack` package and overviews common data input and analysis pipelines. For a tutorial about creating custom network functions and network statistics see the "Custom Functions in `netjack`" vignette.

##Introduction

Samples of _registered_ networks, or networks that consist of the same node set, are increasing common in a variety of scientific fields. The `netjack` package implements an framework to let researchers quickly manipulate and analyze large samples of registered networks, as well as develop custom functionality that builds on the existing `netjack` framework. 

In this vignette, we go over the following procedures:

- Basic data objects and function classes in `netjack`
- Data Input
- Network Manipulation Functions
- Network Statistic Functions
- Difference, Group and Group Difference Testing.


##Basic Data Objects and Function Classes

`netjack` is built around a series of S4 classes that represent different levels of network manipulation, and functions that act on each level of network manipulation. This section describes these classes and functions at a summary level.

The most basic data object is the `Net` object. This represents a single network, along with node level variables, such as partition assignments. 

A `NetSample` object represents a collection of `Net` objects, along with network level variables. For example, if each network in the sample represents a single individual's functional brain network, a network level variable could be the diagnostic status of each individual. 

The `Net` and `NetSample` objects are representations of raw data. To work with these data objects, the `net_apply` function can be used to apply a _network manipulation function_. This class of functions take a single network, and perform a series of manipulations on the network, returning the manipulated networks. As an example the `node_jackknife()` function applied to a `Net` object returns a set of `Net` objects corresponding to the original network with each node removed in turn.

To represent the output of a `net_apply`, the S4 classes `NetSet` and `NetSampleSet` are used. These classes represent both the original `Net` or `NetSample`, as well as the product of the `net_apply`.

One common procedure that network analysis uses is the calculation of various network statistics. In `netjack` _network statistic functions_ can be used via the `net_stat_apply` function to quickly be calculated for both `NetSet` and `NetSampleSet` objects.  The output of a `net_stat_apply` is a `NetStatSet` or `NetSampleStatSet` object. 

`netjack` implements several statistical testing procedures that are described below. Additionally, to extract a `data.frame` of the calculated network statistics from a `NetStatSet` or `NetSampleStatSet` object, `to_data_frame()` can be used.

###Data Input

To illustrate the various features of `netjack`, two simulated datasets are provided, GroupA and GroupB.  Networks can be loaded into the `netjack` framework from adjacency matrices, either as single Net objects, or more commonly as one NetSample object.

```{r}
library(netjack)
data("GroupA")

Subject1 <- as_Net(GroupA[[1]], "Subject1")
show(Subject1)

```

Node Variables can be assigned during construction as a named list:

```{r}
Subject2 <- as_Net(GroupA[[2]], "Subject1", node.variables = list(community = c(rep(1,10), rep(2,10))))

show(Subject2)

```

Typically, a researcher using `netjack` is analyzing a sample of registered networks rather than a single network. `NetSample` objects can be constructed in much the same way as a `Net` object can, using lists of adjacency matrices rather than a single matrix:

```{r}
GroupASamp = as_NetSample(GroupA, net.names = as.character(1:20) , node.variables = list(community = c(rep(1,10), rep(2,10))), sample.variables = list(group = rep(1, 20)))

show(GroupASamp)


```

Importantly, when a NetSample object is created, the list of node variables is applied to every network. This is appropriate in registered network applications, where for example, in neuroimaging networks, each node represents a specific brain region, and each node is the same for each subject.

Sample variables represent network level characteristics. For example, if each network represents a functional connectivity network from a neuroimaging study, a sample variable might be the diagnostic status of a particular individual. 

###Network Manipulation Functions

Once a sample of networks is represented as a `NetSample` object, a network manipulation function can be applied. As described previously, these functions change a network in some way. As an example, the `node_jackknife` function returns a set of networks, where each node has been removed in turn.

Network manipulation functions can be applied via `net_apply` to a `Net` object to produce a `NetSet`, or can be applied via `net_apply` to a `NetSample` object to produce a `NetSampleSet`

```{r}
Sub1Jackknifed <- net_apply(network = Subject1, net.function = "node_jackknife")

show(Sub1Jackknifed)

GroupAJackknifed <- net_apply(network = GroupASamp, net.function = "node_jackknife")

show(GroupAJackknifed)
```

Network manipulation functions that involve node level variables can be used by including them in the `net.function.args` argument within `net_apply`. For example, `network_jackknife` removes sub-networks on the basis of a node level grouping variable.

```{r}

GroupANetJackknifed <- net_apply(GroupASamp, net.function = "network_jackknife", net.function.args = list(network.variable = "community"))

show(GroupANetJackknifed)

```

Once a network manipulation function has been applied, network statistics can be computed.

###Network Statistic Functions

A network statistic is a single numerical summary of some aspect of a network's structure or topology. `netjack` focuses on the analysis of networks at a network statistic level, and provides simple interfaces for calculating network statistics on collections of networks.

Similar in structure to the network manipulation functions, network statistic functions are applied via a `net_stat_apply` function, which can be used with either a `NetSet` object, or a `NetSampleSet` object. This produces a `NetStatSet` and a `NetSampleStatSet` respectively.

```{r}

Sub1JackknifedGlobEff <- net_stat_apply(Sub1Jackknifed, net.stat.fun = global_efficiency)

show(Sub1JackknifedGlobEff)

GroupAJackknifedGlobEff <- net_stat_apply(GroupAJackknifed, net.stat.fun = global_efficiency)

show(GroupAJackknifedGlobEff)


```

Once a `NetStatSet` or `NetSampleStatSet` has been computed, the computed network statistics can be extracted into a `data.frame` by using the `to_data_frame` function. The data frame returned is in long format, with a row for each manipulated network.

```{r}

Sub1Data = to_data_frame(Sub1JackknifedGlobEff)

names(Sub1Data)

GroupAData = to_data_frame(GroupAJackknifedGlobEff)

head(GroupAData)

```

###Difference, Group and Group Difference Testing.

`netjack` implements three statistical testing procedures in easy to use functions for both tabular and graphical output. The first test is the _difference test_ which assess if any specific network manipulation causes a significant difference from the original network in a given network statistic. This test is implemented with the `diff_test` and graphically with the `net_ggPlot` function. Plotting uses the `ggplot` package, making the aesthetic presentation easily manipulated.

The example dataset `GroupA` has been generated so that the removal of node 10 will result in a significant difference in global efficiency from the original networks. Below are the full set of steps for this analysis:

```{r fig.height=5, fig.width=6, fig.align="center"}

GroupASamp = as_NetSample(GroupA, net.names = as.character(1:20))
GroupAJackknifed = net_apply(GroupASamp, net.function = "node_jackknife")
GroupAJackknifedGlobEff = net_stat_apply(GroupAJackknifed, net.stat.fun =  "global_efficiency")

diff_test(GroupAJackknifedGlobEff)
net_ggPlot(GroupAJackknifedGlobEff)


```

The second test implemented is the group test. This examines differences between to sample level groups (such as healthy controls and individuals with a disorder) in a network statistic, subject to a network manipulation.

In this example, `GroupA` has been simulated to have node 10 be important for global efficiency, while `GroupB` has node 15 as important for global efficiency. We combine these datasets into a single object, and perform the group testing now.

```{r fig.height=5, fig.width=6, fig.align="center"}

fullGroup = c(GroupA, GroupB)

fullSamp = as_NetSample(fullGroup,net.names = as.character(1:40), sample.variables = list(group = c(rep("GroupA", 20), rep("GroupB", 20))))

fullSampJackknifed = net_apply(fullSamp, net.function = "node_jackknife")
fullSampleJackknifedGlobEff = net_stat_apply(fullSampJackknifed, net.stat.fun = "global_efficiency")

group_test(fullSampleJackknifedGlobEff, grouping.variable = "group")

netGroup_ggPlot(fullSampleJackknifedGlobEff, grouping.variable="group")

```

Finally, the group difference test assesses if the network manipulation has a differential impact on the network statistic between the groups. This test is implemented with `group_diff_test` and graphically with `netGroupDiff_ggPlot`.

```{r fig.height=5, fig.width=6, fig.align="center"}
group_diff_test(fullSampleJackknifedGlobEff, grouping.variable = "group")

netGroupDiff_ggPlot(fullSampleJackknifedGlobEff, grouping.variable="group")

```

From this, we can see that when node 10 or node 15 are removed, this results in a significantly different change from the original global efficiency value between Group A and Group B.

